The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2010-12-03 smurf_paper.tex
description for later PCA modes, and intro to SMURF algorithm section

@@

>>>>

The subsequent components identified by the PCA analysis in
Fig.~\ref{fig:pca} vary wildly. Component 3, though not obvious from
its eigenvector plot, is almost a pure ramping waveform that only
appears in the s4a data, and produces the broad line featuress
centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
Fig.~\ref{fig:psd}. In other data sets, this third component has also
often instead exhibited a regularly-spaced series of spikes with
heights quantized into several amplitude families (again, only in the
s4a array). In both cases, the shape and phase of this signal is the
same in each bolometer, although the amplitude (and sign) varies, as
evidenced by the large random scatter in the s4a eigenvalue maps. It
is thought that these features are some sort of data acquisition
glitch as the spikes (when they appear) exhibit the expected point
response function of the anti-aliasing filter. Similarly, component 4
also appears to be a data acquisition glitch in the s4a array,
although along a single column, and with a peak at a lower frequency
of 40\,Hz, and a second feature right at the Nyquist frequency.

The last two components in Fig.~\ref{fig:pca} have an entirely
different character. The amplitudes of both components are greatest in
low-responsivity regions of the two subarrays (lower-right at
450\,\micron\ and bottom-left at 850\,\micron). Bolometers in these
regions of the subarray are probably only tenuously locked, and are
more sensitive to changes in the loading, possibly resulting in the
step and spikes that are seen in component 5. Component 6 is clearly
related to component 5, though with the absence of the steps and
spikes, so that together they seem to represent a continuum of shapes
that are linear combinations of the spikes and steps with the slower
drift evident only in component 6.

@@

Given predictions for the performance of a fully-functioning SCUBA-2,
and the observed properties of the first-installed subarrays during
commissioning in late 2009 as described in the previous section, SMURF
was designed with two primary goals in mind: (i) reduced maps should
be scientifically useful and of ``near publication quality'' with
minimal user interaction; and (ii) the execution time should not
appreciably exceed the amount of time required to conduct
observations. By attaining these goals it would be possible to conduct
real-time reduction at the telescope to offer feedback to observers,
and it would also be possible to maintain a central repository of
(nearly) publishable data products.

There are two general families of existing map-making strategies that
were considered. The first are \emph{maximum likelihood} methods that
were originally pioneered for the analysis of Cosmic Microwave
Background (CMB) maps. Here, the pixels in the map are considered to
be free parameters in a model-fitting problem, and the data are the
raw time-series. By estimating the full noise covariance matrix for
the data, it is then possible to estimate the maximum-likelihood
map. While direct inversion solutions to this problem are untenable,
some recent efforts to optimize these algorithms have enabled the
reduction of smaller data sets in a reasonable amount of time. While
maximum likelihood maps are, theoretically, the best maps that can be
made from a noise point of view, the required computing resources and
execution time would not have been feasible for the proposed real-time
data reduction and data archiving model.
